---
title: "BSPR final assessment"
author: "Haiqing Zhong"
date: "`r date()`"
format: 
  html:
    fig-width: 8
    fig-height: 4
    code-fold: true
---

# Introduction

This the final assessment of Basic Statistics and Projects in R course. We will explore two data set in the package medicaldata to find something interesting.

# Preparation

Firstly, we should install R packages and load them to R studio.

```{r}
#| output: false
#| warning: false
# Set language
Sys.setenv(LANG = "en")

# Clear memory
rm(list=ls())
gc()

# Load libraries
library_names <- c("tidyverse", "knitr", "downlit", "ggcorrplot","medicaldata", "rstatix", "datarium","PairedData")

lapply(library_names, function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x)
    library(x, character.only = TRUE)
  }
})

# loading packages #
library(tidyverse)
library(ggplot2)
library(here)
library(unibeCols)
library(medicaldata)
library(knitr)
library(snapshot)
library(bain)
library(ggpubr)
```

# Data

## Explore the dataset covid_testing in Medicaldata

### Read and process data

1, We read and process data of COVID-19 testing at CHOP in 2020 which is from the dataset Medicaldata.

```{r}
#| warning: false
# read data #
data_covid_testing <- medicaldata::covid_testing

str(data_covid_testing)

# display the first 5 rows of the data set
head(data_covid_testing,5)
```

### Mean and median calculation

2, We want to explore if there is difference in the pan_days between man and woman

```{r}
#| warning: false
# calculate the mean, median, standard deviation of ct value #
summary_stat <- data_covid_testing %>%
  group_by(gender) %>%                        
  summarise(Mean_pan = mean(pan_day, na.rm = TRUE),
            Median_pan = median(pan_day, na.rm = TRUE),
            Variance_pan = var(pan_day, na.rm = TRUE),
            SD_ct = sd(pan_day, na.rm = TRUE),
            IQR_ct = IQR(pan_day, na.rm = TRUE))

summary_stat %>%
  kable(digits = 2)
```

### Box plot and scatter plot

3, box plot and scatter plot

```{r}
# creat boxplot and scatter plot #

data_covid_testing %>%
  ggplot(aes(x = gender, y = pan_day, fill = gender)) +
  geom_boxplot() + 
  theme_bw()

data_covid_testing %>%
  ggplot(aes(x = age, y = pan_day, fill = gender, colour = gender)) +
  geom_point() + 
  theme_bw()
```

### Two sample t test

4, then we do a t-test to see if there is a significant difference between female and male

```{r}
#| warning: false
data_covid_testing %>%
  ggplot(aes(x = gender, pan_day, fill = gender)) +
  geom_boxplot() +
  theme_bw() +
  stat_compare_means(aes(label = after_stat(p.signif)), method = "t.test", ref.group = "female")
```

### Conclusion

5, The p value is above 0.5, there is no significant difference in the pan_days between man and woman.

## Explore the dataset smartplill in medicaldata

we want to see the distribution of the age

```{r}
#| warning: false
smartpill |> 
  ggplot(aes(Age)) +
  geom_histogram(bins = 20,
                 color = "black",
                 fill = "steelblue") + 
  theme_bw() + 
  geom_vline(aes(xintercept=mean(smartpill$Age, na.rm =  TRUE)), color="red") +
  geom_vline(aes(xintercept=median(smartpill$Age, na.rm =  TRUE)), color="blue")
```

Generate a quantile-quantile (QQ) plot using `geom_qq` and `geom_qq_line` to assess its alignment with the theoretical line.

```{r}
#| warning: false
smartpill %>%
  ggplot(aes(sample = Age)) + 
  geom_qq_line(distribution = stats::qnorm) +
  geom_qq(color = "steelblue", distribution = stats::qnorm) + 
  xlab("Theoretical Quantiles") +
  ylab("Sample Quantiles") +
  theme_bw()
```

then we test if it's normal distribution

```{r}
#| warning: false
smartpill %>%
  shapiro_test(Age) %>%
  kable(digits = 10)
```

conclusion: p value is below 0.5, so the age is non-normal distributed.
